\chapter*{Macro Lecture II.4: Fiscal and Monetary Policy -- The IS-MP (IS-LM) Model}
\section*{Overview}
This lecture integrates fiscal and monetary policy into the general equilibrium framework of the goods and money markets using the IS-MP (IS-LM) model. With fixed prices, the model explains how output and interest rates are jointly determined by both fiscal actions and monetary policy.

\section*{The IS Curve: Goods Market Equilibrium}
\begin{itemize}
    \item \textbf{Goods Market Equilibrium:} The equilibrium condition is given by
    \[
    Y = C + I + G + NX.
    \]
    \item \textbf{Modified Consumption and Investment Functions:}
    \begin{itemize}
        \item Consumption: 
        \[
        C = A_0 - a \cdot i + c(1-t)Y.
        \]
        \item Investment:
        \[
        I = I_0 + \text{mpi} \cdot Y - b \cdot i.
        \]
    \end{itemize}
    \item \textbf{Derivation of the IS Curve:} Combining the above functions, the equilibrium condition becomes:
    \[
    Y\Bigl(1 - c(1-t) - \text{mpi} + z\Bigr) = \Bigl[A_0 + I_0 + G + X\Bigr] - (a+b)i,
    \]
    which can be rearranged to express output \(Y\) as a function of the interest rate \(i\).
\end{itemize}

\section*{The MP Curve: Monetary Policy Rule}
\begin{itemize}
    \item The MP curve represents combinations of output \(Y\) and the interest rate \(i\) that satisfy money market equilibrium.
    \item Under the fixed price assumption, nominal and real interest rates coincide.
    \item The Central Bank adjusts \(i\) to stabilize output---lowering \(i\) when output is below potential and raising \(i\) when output exceeds potential.
\end{itemize}

\section*{General Equilibrium and Policy Analysis}
\begin{itemize}
    \item \textbf{Equilibrium:} General equilibrium is achieved where the IS and MP curves intersect, determining the equilibrium output \(Y^*\) and interest rate \(i^*\).
    \item \textbf{Policy Implications:}
    \begin{itemize}
        \item \emph{Fiscal Policy:} Changes in government spending \(G\) or taxation \(T\) shift the IS curve.
        \item \emph{Monetary Policy:} Adjustments to the interest rate target shift the MP curve.
        \item \emph{Crowding Out:} The degree to which government spending crowds out private investment depends on the slopes of the IS and MP curves.
    \end{itemize}
    \item \textbf{Analysis Steps:}
    \begin{enumerate}
        \item Identify shifts in the IS and/or MP curves.
        \item Determine the new equilibrium \((Y^*, i^*)\).
        \item Discuss the resulting effects on both the goods and money markets.
    \end{enumerate}
\end{itemize}